Structured Versus Decorated Cospans

### 2022-03-17 ResponseToReviews.tex

\section{Our Answer}

\begin{enumerate}

\item As the only point against, I feel there is a growing number of “cospan” articles. The present work seems like it could serve as a 
comparison of 
all of 
these, but perhaps that comparison could be more explicit and detailed: what is it done in each one of these “cospan” articles? is there anything 
that 
is not already explained in this one? If some of the articles about decorated cospans were found to contain errors, is this the only construction of 
decorated cospans that solves these errors?

{\bf The present work reviews all known previous work on decorated and structured cospans in Section 6, and we now point this out in the end of 
the Introduction.   As we mention, our previous paper ``Structured 
cospans'' fixed the errors in previous work on decorated cospans by using structured cospans instead.   Here we describe how the new improved 
decorated cospans can also fix those errors.     There may be even more ways to correct these errors, but we do not know them.}

{\kenny $\checkmark$}  {\john $\checkmark$} {\chris $\checkmark$}

\iffalse
{\chris As was suggested by John at some point, would we perhaps want to add a sentence or two in the introduction, saying that indeed there are a 
number of cospan articles etc. and that in the beginning of section 6 we elaborate on those?}

{\kenny I'm fine either way on this.}

{\john I have added a remark in the Introduction, ``This result [Theorem 4.1] shows that under certain conditions, structured and decorated cospans 
provide equivalent ways of describing open systems.  We illustrate this in Section 6 with applications to graphs, electrical circuits, Markov 
processes, Petri nets, Petri nets with rates, and dynamical systems.   This is meant to be a fairly thorough review of the existing literature.''

If you're happy with my response above give me a check mark, Christina!}

{\chris Looks good, I only added ``in the end of'' the introduction, if it is OK, John, erase current comment.}
\fi

@@

 \item Lemma 2.4: when talking about the functors $F : (\mathsf{A},\otimes)\to(\mathsf{Cat},\times)$, I think sometimes it is specified that they are lax and sometimes 
it is not. Perhaps it would be clearer to always use “lax monoidal” or to just use “monoidal” to mean lax monoidal.

{\bf We have added ``lax'' at Lemma 2.4 in two occurences, and have checked the rest of the document to verify that this does not happen 
elsewhere.}

@@

{\bf There is now a paper detailing how this software works, so according to this suggestion we have changed our paper so it simply refers to that other paper.}   {\john But we still need to add an arXiv link in reference 1!  It should be on the arXiv soon.}

{\kenny Can we put ``In preparation." if they don't have this on the arXiv in a timely manner?}

{\chris Sounds good to me...}

{\john Yes, we can do that.  Their paper is written, but they are struggling to get it onto the arXiv... and having trouble for stupid technical reasons.   I'll see if they succeed by the time we submit this.}

@@

%{\chris Great! The 
%hyperref link ??? opens up notes by Kenny for me, I don't know how or why.}

%{\john It's weird that the line ??? opens up anything for you.  It'll go away when we get the arXiv number.} Indeed!

\item Section 6: as a reader, I really appreciate the examples: are there more in the literature? I do not think it is necessary to go in detail over each one of them, but perhaps it would be useful to point to other places where they are described.

{\bf We have added a word about Markov processes in Section 6.2, and with this, all the applications of structured or decorated cospans that we know 
of are discussed here.}

@@

\vskip 1em

{\chris Second review starts here}
\item Given the title and abstract of this paper “Structured vs Decorated Cospans” I would expect this paper to provide a judgment that structured  cospans are strictly better than decorated cospans in all circumstances and a justification for that claim, or a detailed analysis of the pros and cons of 
structured vs decorated cospans, some case studies in applications that show how the difference between these double categories is felt in the 
application domain, and a some description of the tradeoffs. This paper should come with some practical advice for the applied category theorist on 
how to choose whether to model their application with structured vs decorated cospans.

{\bf In fact our goal, as explained in the abstract and introduction, was to prove that the structured and decorated cospan approaches are equivalent under certain conditions, and to illustrate this result with a wide variety of examples.  These examples can serve as a guide for other similar applications, while new applications may require exploring which approach works best.  We do not make a stand that either is strictly better.   However, we have changed the end of the Introduction to read as follows:

\vskip 1em
\begin{quote}
This result shows that under certain conditions, structured and decorated cospans provide equivalent ways of describing open systems.  We illustrate this in Section 6 with applications to graphs, electrical circuits, Markov processes, Petri nets, Petri nets with rates, and dynamical systems.   This is meant to be a fairly thorough review of the existing literature.   It becomes clear that when either structured or decorated cospans can be used, structured cospans are simpler.  However, for open dynamical systems we need decorated cospans, for reasons we explain.   Thus, to describe a map from open Petri nets with rates to open dynamical systems, we use decorated cospans.
\end{quote}
}

@@

{\kenny The boldface looks good to me, but I also like Christina's suggestions above if we really want to make an effort to appease this referee. I'm fine either way on this.}

\fi

{\john See if you like my new remarks in boldface, and give me a check mark if you do!  Kenny had checked this one off already, but I've changed it a 
lot.}

{\chris Looks great to me. Final two minor points, feel free to ignore or not and delete this comment. (1) Is ``apparently'' supposed to sound a bit 
sceptical? If not, would ``actually'' be better or not?}

{\john I said ``apparently'' because two referees suggested that we might be able to create a structured cospan category of open dynamical systems if we could find more morphisms between dynamical systems.  But since the readers of our paper aren't being told about this hope, I guess it will confuse \emph{them} if we say ``apparently''.   So I've deleted ``apparently'' in this sentence:

\begin{quote}
However, for open dynamical systems we apparently need decorated cospans, for reasons we explain. 
\end{quote}
\noindent
both here and in the actual paper.  The referees may complain, but the readers will be less puzzled.}

{\chris (2) If we were being really accurate, don't we 
actually ONLY use decorated cospans to describe the map from open Petri nets with rates to open dynamical systems? Namely since dynamical systems only 
work as decorated cospans, we are required to use decorated cospans also for Petri nets to then get the map between them. Unless I am mixing something 
there.}

{\john You're right.  So I've also changed this sentence above and in the paper:

\begin{quote}
Thus, to describe a map from open Petri nets with rates to open dynamical systems, we use \emph{both} structured and decorated cospans.
\end{quote}}

\item Given that Petri nets with structured cospans works, and Reaction Networks work, but open dynamical systems do not, I would like some more 
explanation of the differences here. The examples of graph, labeled 
graph, Petri net, reaction net are all examples of attributed C-Sets [1], the structured cospan construction in provided for arbitrary attributed 
C-Sets in Catlab.CategoricalAlgebra [2]. The open dynamical systems are not attributed C-Sets.(...)The examples in section 6 would be much stronger 
if the others included a theorem that all (or finitary) categories of presheaves satisfy the 
conditions of theorem 4.1 or a counter example (...) Without this analysis, I am left with the intuition
that for discrete structures like finitary presheaves over a finitely presented schema, the decorated and structured cospans are going
to be equivalent (...) If that intuition is correct, the paper should say something about it. If the intuition is incorrect, then additional 
explanation should be provided. Hopefully it is easy to check and provide some commentary on it in this paper.

{\bf Unfortunately not all cases of C-sets meet the conditions of Theorem 4.1.   For example, the usual forgetful functor from $\mathsf{Graph}$ to $\mathsf{Set}$ sending any graph to its set of vertices is an opfibration, but not the forgetful functor sending any graph to its set of edges.  So, this gives a structured cospan category to which Theorem 4.1 does not apply.   Finding extra conditions such that it \emph{does} apply is an interesting question, but it would take more thought.}

@@

{\john I've improved the reference and added a citation to it.  Is this okay now?}

{\chris Yes!}
\fi

%[Several points in the text are written in a way style that can be interpreted as the dismissive style common to many discussions of Category 
%Theory. I would recommend rewriting at least the following points that should be made in a more collegial tone.]

\item pg. 2: ``But this hope is doomed for reasons painstakingly explained in [3, Section 5] … despite all the tricks one might try.'' This paragraph 
provides 
no insight into the reasons that those tricks fail. Referring them to tricks is dismissive. The citation is to a paper in TAC by the current authors.

{\bf We removed the part of the sentence ``despite all the tricks one might try'' according to this suggestion. In [3, Section 5] we spent five pages discussing these attempts and why they fail.    The diligent reader may wish to read this discussion and see if there is some other way out.}

 {\chris \checkmark} {\john \checkmark}

\iffalse
{\john He thought we were being dismissive because he didn't know that the ``tricks'' were all things I'd done myself!  If he looked at [1, Section 5] 
he'd see it was me coming up with a bunch of tricks to try to save the decorated cospan category of open graphs, and then Kenny and me explaining why 
this tricks don't work.  So maybe I should just delete ``despite all the tricks one might try''.  Sound good?}

@@

{\chris Sounds good to me! See above, amend at will.}

{\kenny Sounds good.}
\fi

\item pg. 3: ``Many of the flawed applications of decorated cospans have been fixed using structured cospans.” This dismisses the work of other 
scholars 
using the decorated cospan approach as flawed. The alleged flaw should be made more precise. What about the decorated cospan approach is inferior to 
the structured cospan approach?''
%The goal of this paper should be to illustrate to an applied category theory audience the differences between 
%decorated and structured cospans. Referring back to previous work of the authors to make that point does not provide value.

{\bf  The flaws---mathematical errors that render a number of claimed theorems incorrect---are now briefly explained in  in the Introduction, in the paragraph beginning ``Unfortunately, many applications of decorated cospans were flawed.''.  In reference [3], Kenny Courser and John Baez analyzed these mistakes in detail and fixed them using structured cospans.  See also comment (9).}

@@

{\chris Right, done. About our answer here, we could politely tell them that the problem is simply that they are wrong! And perhaps, in 
text amend that sentence to ``Many of the flawed applications of decorated cospans, \emph{like the open graph one described above},...'' because we 
do actually say what the flaw is there, correct? I don't see where we describe flaws in Section 6.}

{\kenny It looks like we loosely mention what the problem is on page 2, but we already explained (in detail) what the problems are in \emph{Structured cospans}. I don't see why we can't just point the reader that wants to see what those problems are to that paper. I don't think we explain any flaws in Section 6; only use both frameworks and compare them.}

{\kenny Actually, we do talk about the flaws a bit more in the middle of page 21 when we're setting up the example of open graphs for Theorem 2.2.}

{\john See if you like my boldface response.  I'm acting like we added an explanation of the flaws to our introduction, so we don't have to say ``Hey, 
buster --- we already explained this.''}
\fi

{\chris I would potentially skip the first sentence of our response above. Does it sound a bit hostile? And perhaps he also means other people that 
may have used this approach, or perhaps Brendan himself (not a coauthor at the time?) that invented them. Not too important in any case.}

{\john Okay, I deleted it.  Indeed, it \emph{was} a bit hostile.}

@@

\item pg. 27: To show that U does not have a left adjoint, we use the following well-known result. For readers of the TAC paper previously cited, 
that might 
be a well known result that does not require a citation, but my understanding of the Compositionality audience assumed that the ACT work was targeted 
readers that are more familiar with the applications and thus cannot be assumed to know such results.

{\bf We have eliminated the term ``well-known'' and added a citation to Emily Riehl's book \textsl{ Category Theory in Context} which includes said result, above Lemma 6.1.}

@@

{\kenny Agreed about keeping the examples together. The boldface looks good to me.}

{\john I tweaked the boldface a tiny bit, so you can revoke your checkmarks if you want; 
otherwise just comment out our little conversation here. By the way, you can comment out
a big chunk of text just by putting $\backslash$iffalse in front of it and $\backslash$fi at the end.  This is a special case of a trick for making some text appear only if some condition is true!}

{\chris Fourth review starts here}
\item Composition of morphisms f, g in this paper is usually denoted gf, and occasionally gof. I’d suggest mentioning this in the “conventions” 
section.

{\bf We added the sentence ``For composition of morphisms we use concatenation or occasionally $\circ$.'' in the convention section.}

@@

%{\john Great.  Except for one thing: I hate footnotes.   If something is worth saying, just say it, don't stick a little number in the text and make 
%the reader decide if it's worth looking down at the bottom of the page to see what this mysterious number means.  That's more distracting than just 
%saying it!   So I moved the comment into the text and merged it into the previous sentence.}

%{\chris OK!}

\item Theorem 2.1: Maybe worth saying explicitly that $s$ is an object of the category $F(m)$ rather than using $\in$, since this is exactly the  distinction to  Fong's original definition.

{\bf  We have added a remark emphasizing this right after Theorem 2.1.}

@@

%{\john It's hard to tweak the statement, since we're really trying to introduce the notation of decorations in maps of structured cospans here!    
%So 
%I've added something about this \emph{after} the statement of Theorem 2.1.  See if you like it:
%\begin{quote} 
%Note that the decoration $s \in F(m)$ is now an object in the category $F(m)$, not an element of a set as in Fong's original approach. 
%\end{quote}}

%{\chris Looks good!}

\item Proof of theorem 2.1: You should probably at least handwave that the cartesian liftings really are cartesian liftings. (No doubt it’s easy, but 
this 
is a classic sort of spot where errors can hide.)

\textbf{Kenny Courser carefully checked that these really are co/cartesian liftings, but the argument is straightforward and uninteresting so we will just 
add remarks saying that this can be shown.}
